### 2022-06-29 sections/introduction.tex
revised to V2.1

%-------------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
%-------------------------------------------------------------------------------
After the discovery of the Higgs boson in 2012 at the Large Hadron Collider (LHC) by the ATLAS~\cite{Aad:2012tfa} and 
CMS~\cite{Chatrchyan:2012ufa} collaborations, 
%Since the observation of Higgs boson by the ATLAS and CMS experiments~\cite{Aad:2012tfa,Chatrchyan:2012ufa} at
a comprehensive programme of measurements % of its properties
has been conducted to explore this particle. Measurements so far have proved to be consistent with the Standard Model (SM) predictions. 
However, the program is ongoing and precision measurements as well as searches for rare new physics processes beyond the Standard Model (BSM)
are underway. One such possibility is flavour-changing neutral-current (FCNC) interactions between the Higgs boson, 
the top quark, and an up-type quark, $tqH$ ($q=u, c$), which have been sought for by the ATLAS and CMS collaborations.  
%with increased precision and proved to be consistent with the Standard Model (SM) predictions. 
%for nearly a decade, by which no significant sign of deviation from Standard Model (SM) was observed until today.
%However the LHC still provides an excellent opportunity to search for rare new physics processes beyond the Standard Model (BSM).  
%there are still chances to find new physics by improving the sensitivity of previous searches.
%The flavour-changing neutral-current (FCNC) interactions 
%between the Higgs boson, the top quark, and a $u$- or $c$-quark, $tqH$ ($q=u,c$) has been actively searched for by the ATLAS and CMS collaborations. %is still an active topic.
Since the Higgs boson is lighter than the top quark~\cite{Aad:2015zhl},
%with a measured mass $m_H=125.09 \pm 0.24~\gev$~\cite{Aad:2015zhl}, 
such interactions could manifest themselves as FCNC top-quark decays~($t\to qH$)~\cite{Agashe:2013hma}.
%Furthermore, if the $tqH$ interaction exists, the associated single-top and  Higgs  production process through this interaction would enhance 
%becomes non-zero, enhancing 
%the total
%production cross section of $pp\rightarrow tH$.

In the Standard Model (SM), the FCNC interaction is forbidden at tree level and suppressed at higher orders due to the Glashow-Iliopoulos-Maiani (GIM) mechanism~\cite{Glashow:1970gm}. An observation of an enhanced rate of this decay would be clear evidence for new physics.

@@

However, these branching ratios can be as high as $\BR(t\to qH) \sim 10^{-5}$ when processes beyond the SM are included. Examples of these processes include: quark-singlet models~\cite{AguilarSaavedra:2002kr}, two-Higgs-doublet models (2HDM)~\cite{ Branco:2hdm2012} with or without flavour violation,

@@

An even larger branching ratio of  $\BR(t\to cH) \sim 10^{-3}$ can be reached in 2HDM without explicit flavour conservation, since in them the tree-level FCNC coupling is no longer forbidden by any symmetry~\cite{Cheng:1987rs,Baum:2008qm,Chen:2013qta,Chiang:2015cba,Crivellin:2015hha,Botella:2015hoa, Gori:2017tvg,Chiang:2017fjr}. 

@@

The $tH$ associated production in the SM is expected to have a cross section of $92^{+7}_{-12}$ fb at a centre-of-mass energy of $\sqrt{s}= 13~\tev$~\cite{deFlorian:2016spz}.

@@

of top-quark pair ($\ttbar$) events collected in proton-proton ($pp$) collisions at centre-of-mass energies of $\sqrt{s}=7~\tev$ and $8~\tev$~\cite{Aad:2014dya,Aad:2015pja,Khachatryan:2016atv} during Run~1 of the LHC, as well as at $\sqrt{s}=13~\tev$~\cite{fcnc36} using early Run~2 data.

@@

 $\BR(t\to qH)$, denoted~\footnote{In the following, $WbHq$ is used to denote both $W^+b H\bar{q}$ and its charge conjugate, $HqW^- \bar{b}$. Similarly, 
$WbWb$ is used to denote $W^+b W^- \bar{b}$.} as $\ttbar \to WbHq$.  The Higgs boson is assumed to have a mass of $m_H=125~\gev$ and to decay as predicted by

@@

Using 36.1~fb$^{-1}$ of data at $\sqrt{s}=13~\tev$, the ATLAS Collaboration has derived upper limits at 95\% confidence level (CL) on the branching ratio
$\BR(t\to cH)<0.22\%$ using $H\to \gamma\gamma$ decays~\cite{Aaboud:2017mfd} and $\BR(t\to cH)<0.16\%$ based on

@@

multilepton (electron or muon) signatures resulting from 

@@

The upper limits are set on $\BR(t\to cH)<0.42\%$ using $H\to b\bar{b}$ decay~\cite{fcnc36} and $\BR(t\to cH)<0.19\%$ using the $H\to \tau^+\tau^-$ decay in which at least

@@

These upper limits are derived assuming that the branching ratio $\BR(t\to uH)$ is null. Similar upper limits are obtained for $\BR(t\to uH)$ assuming $\BR(t\to cH)=0$.

@@

The CMS Collaboration performed a similar search using  
$H\to b\bar{b}$ decays~\cite{Sirunyan:2017uae} with 35.9 fb$^{-1}$ of data at $\sqrt{s}=13~\tev$, resulting 
in upper limits of $\BR(t\to cH)<0.47\%$ and $\BR(t\to uH)<0.47\%$, in each case neglecting the other decay mode.
%Compared with previous searches, 
The search in Ref.~\cite{Sirunyan:2017uae} considers in addition the contribution to the signal from 
$pp \to tH$ production~\cite{Greljo:2014dka}. 
CMS subsequently updated their search for $tqH$ in the $H\to b\bar{b}$ channel using
the full Run 2 dataset; observing (expected) upper limits on  $\BR(t\to cH)$ of $9.4\times10^{-4}\, (8.6\times10^{-4})$
and on $\BR{B}(t\to uH)$ of $7.9\times10^{-4}\, (1.1\times10^{-3})$~\cite{CMS:2021gfa}.
%CMS also recently reported a search for $tqH$ in the $H\rightarrow \gamma\gamma$ mode using the full Run-2 data and set observed(expected) upper limits on $\mathcal{B}(t\to Hu)$ of $1.9\times10^{-4}(3.1\times10^{-4})$ and $\mathcal{B}(t\to Hc)$ of $7.3\times10^{-4}(5.1\times10^{-4})$~\cite{CMS-PAS-TOP-20-007}.

\begin{table}[t!]
\caption{\small{Summary of 95\% CL upper limits on $\BR(t \to cH)$ and $\BR(t \to uH)$ obtained by the ATLAS and CMS Collaborations with Run 2 data. Each limit is obtained assuming the other branching ratio is null.}}
\begin{center}
\small 
\begin{tabular}{ccccc}
\toprule\toprule
& &\multirow{2}{*}{$\mathcal{L}$ [fb$^{-1}$]} & \multicolumn{2}{c}{95\% CL observed upper limits}  \\
& & 										    & \multicolumn{1}{c}{on $\BR(t \to cH)$}            & \multicolumn{1}{c}{on $\BR(t \to uH)$} \\
\midrule
\multirow{5}{*}{ATLAS}
& $H \to b\bar{b}$~\cite{fcnc36}                                          & 36.1         & $4.2 \times 10^{-3}$ & $5.2 \times 10^{-3}$ \\
& $H \to \gamma\gamma$~\cite{Aaboud:2017mfd}                              & 36.1         & $2.2 \times 10^{-3}$  & $2.4 \times 10^{-3}$  \\
& $H \to \tau\tau$ ($\lephad$, $\hadhad$)~\cite{fcnc36}                   & 36.1         & $1.9 \times 10^{-3}$  & $1.7 \times 10^{-3}$  \\ 
& $H \to WW^*, \tau\tau, ZZ^*$ ($2\ell$SS, $3\ell$)~\cite{Aaboud:2018pob} & 36.1         & $1.6 \times 10^{-3}$  & $1.9 \times 10^{-3}$\\ 
& Combination~\cite{fcnc36}                                               & 36.1         & $1.1 \times 10^{-3}$  & $1.2 \times 10^{-3}$  \\\midrule
\multirow{2}{*}{CMS} 
& $H \to b\bar{b}$~\cite{Sirunyan:2017uae}                                & 35.9         & $4.7 \times 10^{-3}$  & $4.7 \times 10^{-3}$  \\
& $H \to b\bar{b}$~\cite{CMS:2021gfa}                                     & 137          & $9.4 \times 10^{-4}$  & $7.9 \times 10^{-4}$  \\
%& $H \to \gamma\gamma$~\cite{CMS-PAS-TOP-20-007}                          & 137          & $7.3 \times 10^{-4}$  & $1.9 \times 10^{-4}$  \\
% 
\bottomrule\bottomrule
\end{tabular}
\label{tab:limits_summary_ref}
\end{center}
\end{table}

@@

%This paper is focused on the tau decay mode of the Higgs boson with the complete dataset collected during the Run~2 from 2016 to 2018.
The analysis reported here targets Higgs boson decays to 
%This paper is focused on the decay mode of the Higgs boson into
$\tau$-leptons in the complete Run~2 dataset collected in 2015-2018. The corresponding integrated luminosity is 139 fb$^{-1}$. Both $\ttbar \to WbHq$ decays and $pp \to tH$ production are sought. The dataset is divided into several final states depending on the production mode and
the  $W$ and tau decays. Top quark pair production events where the $W$ boson decays hadronically (leptonically) are referred to as $t_h$ ($t_{\ell}$).
The decays $\tau \to \ell \nu_{\ell} \nu_{\tau}$ are referred as $\lep$, while the decays $\tau \to \text{hadrons} + \nu_{\tau}$ are referred as $\thad$.
The contribution of 
$W\rightarrow\tau\nu$ is included as $\lep$ in $t_{\ell}$ when the $\tau$ decays into a light lepton (electron or muon) or as $\thad$ 
in $t_h$ when the $\tau$ decays hadronically.
%either hadronically or leptonically from the $t\rightarrow Wb$ decay, which are referred as $t_h$ and $t_l$. The contribution of $W\rightarrow\tau\nu$ is included either in
%$t_l$ when the $\tau$ decays into a light lepton (electron or muon) as $\tau_{\text{lep}}$ or in $t_h$ when the $\tau$ decays hadronically as $\thad$.
The $H\rightarrow \tau\tau$ decay is detected in
either $\tlhad$ or $\thadhad$ final states when the top decays hadronically ($t_h$), but only the $\thadhad$ final state is considered
when the top decays leptonically ($t_{\ell}$), in order to avoid overlaps with
other ATLAS searches~\cite{Aaboud:2018pob}. 
Events with two hadronically decaying $\tau$-leptons without an electron or muon are named hadronic channels. Events with at least one \tauhad with an additional electron or muon (corresponding to $t_{\ell}$ or $\lep$ events) are named leptonic channels. More signal regions (with the top quark decaying leptonically) are exploited here than in the previous FCNC $\Htautau$ search which was conducted using the partial Run~2 dataset~\cite{fcnc36}.
In addition, an improved treatment of misidentified $\tau$-leptons (fake) in simulation and data-driven
fake estimations from multi-jet background is implemented.
% and more
%$$t\bar{t}$ control regions (CRtt) are used for calibration of misidentified (fake) taus in Monte Carlo and data-driven fake estimations from QCD multi-jet background.
Finally, a multivariate technique based on boosted decision trees is used to discriminate between the signal and the background on the basis of their different kinematical distributions.

>>>>

%-------------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
%-------------------------------------------------------------------------------
Since the discovery of the Higgs boson in 2012 at the Large Hadron Collider (LHC) by the ATLAS~\cite{Aad:2012tfa} and 
CMS~\cite{Chatrchyan:2012ufa} collaborations, 
%Since the observation of Higgs boson by the ATLAS and CMS experiments~\cite{Aad:2012tfa,Chatrchyan:2012ufa} at
a comprehensive programme of measurements % of its properties
has been conducted to explore this particle. Measurements so far have proved to be consistent with the Standard Model (SM) predictions. 
However, the programme is ongoing and precision measurements as well as searches for rare new-physics processes beyond the Standard Model (BSM)
are underway. One such possibility is flavour-changing neutral current (FCNC) interactions between the Higgs boson, 
the top quark, and an up-type quark, $tqH$ ($q=u, c$), which have been searched for by the ATLAS and CMS collaborations.  
%with increased precision and proved to be consistent with the Standard Model (SM) predictions. 
%for nearly a decade, by which no significant sign of deviation from Standard Model (SM) was observed until today.
%However the LHC still provides an excellent opportunity to search for rare new physics processes beyond the Standard Model (BSM).  
%there are still chances to find new physics by improving the sensitivity of previous searches.
%The flavour-changing neutral-current (FCNC) interactions 
%between the Higgs boson, the top quark, and a $u$- or $c$-quark, $tqH$ ($q=u,c$) has been actively searched for by the ATLAS and CMS collaborations. %is still an active topic.
Since the Higgs boson is lighter than the top quark~\cite{Aad:2015zhl},
%with a measured mass $m_H=125.09 \pm 0.24~\gev$~\cite{Aad:2015zhl}, 
such interactions could manifest themselves as FCNC top-quark decays~($t\to qH$)~\cite{Agashe:2013hma}.
%Furthermore, if the $tqH$ interaction exists, the associated single-top and  Higgs  production process through this interaction would enhance 
%becomes non-zero, enhancing 
%the total
%production cross section of $pp\rightarrow tH$.

In the Standard Model (SM), the FCNC interaction is forbidden at tree level and suppressed at higher orders through the Glashow--Iliopoulos--Maiani (GIM) mechanism~\cite{Glashow:1970gm}. An observation of an enhanced rate of this decay would be clear evidence of new physics.

@@

However, these branching ratios can be as high as $\BR(t\to qH) \sim 10^{-5}$ when processes beyond the SM are included. Examples of these processes include: quark-singlet models~\cite{AguilarSaavedra:2002kr}, two-Higgs-doublet models (2HDMs)~\cite{ Branco:2hdm2012} with or without flavour violation,

@@

An even larger branching ratio of  $\BR(t\to cH) \sim 10^{-3}$ can be reached in 2HDMs without explicit flavour conservation, since in these models the tree-level FCNC coupling is no longer forbidden by any symmetry~\cite{Cheng:1987rs,Baum:2008qm,Chen:2013qta,Chiang:2015cba,Crivellin:2015hha,Botella:2015hoa, Gori:2017tvg,Chiang:2017fjr}. 

@@

In the SM, associated production of $tH$ in $pp$ collisions is expected to have a cross section of $\sigma_{tH}=92^{+7}_{-12}$ fb at a centre-of-mass energy of $\sqrt{s}= 13~\tev$~\cite{deFlorian:2016spz}.

@@

of top-quark pair ($\ttbar$) events collected in proton--proton ($pp$) collisions at centre-of-mass energies of $\sqrt{s}=7~\tev$ and $8~\tev$~\cite{Aad:2014dya,Aad:2015pja,Khachatryan:2016atv} during Run~1 of the LHC, as well as at $\sqrt{s}=13~\tev$~\cite{fcnc36} using early Run~2 data.

@@

 $\BR(t\to qH)$, a process denoted by $\ttbar \to WbHq$.\footnote{In the following, $WbHq$ is used to denote both $W^+b H\bar{q}$ and its charge conjugate, $HqW^- \bar{b}$. Similarly, 
$WbWb$ is used to denote $W^+b W^- \bar{b}$.}  The Higgs boson is assumed to have a mass of $m_H=125~\gev$ and to decay as predicted by

@@

Using 36.1~fb$^{-1}$ of data at $\sqrt{s}=13~\tev$, the ATLAS Collaboration has derived upper limits at 95\% confidence level (CL) on the 
$t\to cH$ branching ratio: $\BR(t\to cH)<0.22\%$ using $H\to \gamma\gamma$ decays~\cite{Aaboud:2017mfd} and $\BR(t\to cH)<0.16\%$ based on

@@

multi-lepton (electron or muon) signatures resulting from 

@@

ATLAS also set upper limits of $\BR(t\to cH)<0.42\%$ using $H\to b\bar{b}$ decay~\cite{fcnc36} and $\BR(t\to cH)<0.19\%$ using $H\to \tau^+\tau^-$ decays in which at least

@@

These upper limits are derived assuming that the branching ratio $\BR(t\to uH)=0$. Similar upper limits are obtained for $\BR(t\to uH)$ when assuming $\BR(t\to cH)=0$.

@@

The CMS Collaboration performed a similar search using  
$H\to b\bar{b}$ decays~\cite{Sirunyan:2017uae} with 35.9 fb$^{-1}$ of data at $\sqrt{s}=13~\tev$, resulting 
in upper limits of $\BR(t\to cH)<0.47\%$ and $\BR(t\to uH)<0.47\%$, in each case neglecting the other decay mode.
%Compared with previous searches, 
The search in Ref.~\cite{Sirunyan:2017uae} also considers the contribution to the signal from 
$pp \to tH$ production~\cite{Greljo:2014dka}. 
CMS subsequently updated their search for $tqH$ in the $H\to b\bar{b}$ channel using
the full Run~2 dataset, obtaining observed (expected) upper limits of  $\BR(t\to cH) < 9.4\times10^{-4}\, (8.6\times10^{-4})$
and $\BR(t\to uH) < 7.9\times10^{-4}\, (1.1\times10^{-3})$~\cite{CMS:2021gfa}.
%CMS also recently reported a search for $tqH$ in the $H\rightarrow \gamma\gamma$ mode using the full Run-2 data and set observed(expected) upper limits on $\mathcal{B}(t\to Hu)$ of $1.9\times10^{-4}(3.1\times10^{-4})$ and $\mathcal{B}(t\to Hc)$ of $7.3\times10^{-4}(5.1\times10^{-4})$~\cite{CMS-PAS-TOP-20-007}.

\begin{table}[t!]
\caption{\small{Summary of 95\% CL upper limits on $\BR(t \to cH)$ and $\BR(t \to uH)$ obtained by the ATLAS and CMS collaborations with Run~2 data. Each limit is obtained assuming the other branching ratio is zero.}}
\begin{center}
\small 
\begin{tabular}{ccccc}
\toprule\toprule
& &\multirow{2}{*}{$\mathcal{L}$ [fb$^{-1}$]} & \multicolumn{2}{c}{95\% CL observed upper limits}  \\
& & 										    & \multicolumn{1}{c}{on $\BR(t \to cH)$}            & \multicolumn{1}{c}{on $\BR(t \to uH)$} \\
\midrule
\multirow{5}{*}{ATLAS}
& $H \to b\bar{b}$~\cite{fcnc36}                                          & 36.1         & $4.2 \times 10^{-3}$ & $5.2 \times 10^{-3}$ \\
& $H \to \gamma\gamma$~\cite{Aaboud:2017mfd}                              & 36.1         & $2.2 \times 10^{-3}$  & $2.4 \times 10^{-3}$  \\
& $H \to \tau\tau$ ($\lephad$, $\hadhad$)~\cite{fcnc36}                   & 36.1         & $1.9 \times 10^{-3}$  & $1.7 \times 10^{-3}$  \\ 
& $H \to WW^*, \tau\tau, ZZ^*$ ($2\ell$SS, $3\ell$)~\cite{Aaboud:2018pob} & 36.1         & $1.6 \times 10^{-3}$  & $1.9 \times 10^{-3}$\\ 
& Combination~\cite{fcnc36}                                               & 36.1         & $1.1 \times 10^{-3}$  & $1.2 \times 10^{-3}$  \\\midrule
\multirow{2}{*}{CMS} 
& $H \to b\bar{b}$~\cite{Sirunyan:2017uae}                                & 35.9         & $4.7 \times 10^{-3}$  & $4.7 \times 10^{-3}$  \\
& $H \to b\bar{b}$~\cite{CMS:2021gfa}                                     & 137~~~~~          & $9.4 \times 10^{-4}$  & $7.9 \times 10^{-4}$  \\
%& $H \to \gamma\gamma$~\cite{CMS-PAS-TOP-20-007}                          & 137          & $7.3 \times 10^{-4}$  & $1.9 \times 10^{-4}$  \\
% 
\bottomrule\bottomrule
\end{tabular}
\label{tab:limits_summary_ref}
\end{center}
\end{table}

@@

%This paper is focused on the tau decay mode of the Higgs boson with the complete dataset collected during the Run~2 from 2016 to 2018.
The analysis reported here targets Higgs boson decays into 
%This paper is focused on the decay mode of the Higgs boson into
$\tau$-leptons in the complete Run~2 dataset collected by ATLAS in 2015--2018. The corresponding integrated luminosity is 139 fb$^{-1}$. Both $\ttbar \to WbHq$ decays and $pp \to tH$ production are sought. The dataset is divided into several final states depending on the production mode and
the  $W$ boson and $\tau$-lepton decays. Top quark pair production events where the $W$ boson decays hadronically (leptonically) are denoted by $t_h$ ($t_{\ell}$).
The decays $\tau \to \ell \nu_{\ell} \nu_{\tau}$ are denoted by $\lep$, while the decays $\tau \to \text{hadrons} + \nu_{\tau}$ are denoted by $\thad$.
The contribution of 
$W\rightarrow\tau\nu$ is included as $\lep$ in $t_{\ell}$ when the $\tau$-lepton decays into a light lepton (electron or muon) or as $\thad$ 
in $t_h$ when the $\tau$-lepton decays hadronically.
%either hadronically or leptonically from the $t\rightarrow Wb$ decay, which are referred as $t_h$ and $t_l$. The contribution of $W\rightarrow\tau\nu$ is included either in
%$t_l$ when the $\tau$ decays into a light lepton (electron or muon) as $\tau_{\text{lep}}$ or in $t_h$ when the $\tau$ decays hadronically as $\thad$.
The $H\rightarrow \tau\tau$ decay is detected in
either the $\tlhad$ or $\thadhad$ final state when the top quark decays hadronically ($t_h$), but only the $\thadhad$ final state is considered
when the top quark decays leptonically ($t_{\ell}$), in order to avoid overlaps with
other ATLAS searches~\cite{Aaboud:2018pob}. 
In this search, events with two hadronically decaying $\tau$-leptons and no electron or muon define the hadronic channel. Events with at least one \tauhad and an additional electron or muon (corresponding to $t_{\ell}$ or $\lep$ events) are assigned to leptonic channels. More signal regions (with the top quark decaying leptonically) are exploited here than in the previous FCNC $\Htautau$ search, which was conducted using the partial Run~2 dataset~\cite{fcnc36}.
In addition, an improved treatment of misidentified $\tau$-leptons (`fakes') in simulation and in data-driven
estimations of fakes from multi-jet background is implemented.
% and more
%$$t\bar{t}$ control regions (CRtt) are used for calibration of misidentified (fake) taus in Monte Carlo and data-driven fake estimations from QCD multi-jet background.
Finally, a multivariate technique based on boosted decision trees is used to discriminate between the signal and the background on the basis of their different kinematical distributions.
